---
title: "Exe. 3"
output:
  github_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```
```{r}
library(ggraph)
library(igraph)

library(arrow)
library(tidyverse)
library(gender)
library(wru)
library(lubridate)

library(ggplot2)
library(gridExtra)
library(grid)

library("gender")
```

## Data

```{r}
app <- read_parquet('/Users/danystefan/Documents/01 McGill University/01 MMA/01 Summer 2022/ORGB 672/Assignments/Exe. 3/672_project_data/app_data_sample.parquet')

edges <- read_csv('/Users/danystefan/Documents/01 McGill University/01 MMA/01 Summer 2022/ORGB 672/Assignments/Exe. 3/672_project_data/edges_sample.csv')
```

## Add gender part

```{r}
# first name
names <- app %>% distinct(examiner_name_first)
# names and gender
names_gender <- names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )
names_gender <- names_gender %>% select(examiner_name_first, gender)

# join
app <- app %>% left_join(names_gender, by='examiner_name_first')
```

## Add race part

```{r}
# last names
sur <- app %>% select(surname = examiner_name_last) %>% distinct()

race <- predict_race(voter.file = sur, surname.only = T) %>% as_tibble()
race <- race %>%
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))

# cleanup
race <- race %>% select(surname, race)

#join
app <- app %>% left_join(race, by=c("examiner_name_last" = "surname"))
```

## Add tenure part

```{r}
# get dates
dates <- app %>% select(examiner_id, filing_date, appl_status_date)
# calculate start and end date
dates <- dates %>% mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))

dates <- dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
  ) %>% 
  filter(year(latest_date)<2018)

# join
app <- app %>% left_join(dates, by="examiner_id")
```

## Set up 2 work groups

```{r}
# get 2 work groups
w1 = app[substr(app$examiner_art_unit, 1, 3) == 163,]
w2 = app[substr(app$examiner_art_unit, 1, 3) == 177,]
```

## Summary

```{r}
summary(w1)
summary(w2)
```

## Distributions for gender in each group

```{r}
a <- ggplot(data=w1, aes(x=factor(gender))) + geom_bar(fill="darkblue")  


b <- ggplot(data=w2, aes(x=factor(gender))) + geom_bar(fill="darkred")

a
b
```
For group w1, the ratio is even. However, for w2, there are definitly more male than females.

## Distributions for race in each group

```{r}
a <- ggplot(data=w1, aes(x=factor(race))) + geom_bar(fill="darkblue")  


b <- ggplot(data=w2, aes(x=factor(race))) + geom_bar(fill="darkred")

a
b
```
In both groups, asians and whites are predominantly the majority compared to other races.

```{r}
a <- ggplot(data=w1, aes(x=race)) +
  geom_bar(aes(y = (..count..)/sum(..count..), fill=gender  ), show.legend=FALSE )  +
  ylab("Proportion")+
  ylim(0,1)+
  xlab("Race")
b <- ggplot(data=w2, aes(x=race)) +
  geom_bar(aes(y = (..count..)/sum(..count..), fill=gender  ) )  +
  ylab("Proportion")+
  ylim(0,1)+
  xlab("Race")
grid.arrange(a,b,ncol=2, widths=c(1,1.4), top=textGrob("Race distribution for work groups 163 & 177 colored by examiner gender"))
```

## Network

```{r}
# limit to 2 groups
aus = distinct(subset(app, select=c(examiner_art_unit, examiner_id)))
aus$wg = substr(aus$examiner_art_unit, 1,3)
aus = aus[aus$wg == 163 | aus$wg == 177,]

tm = merge(x=edges, y=aus, by.x="ego_examiner_id", by.y="examiner_id", all.x=TRUE)
tm = tm %>% rename(ego_art_unit=examiner_art_unit, ego_wg=wg)
tm = drop_na(tm)

tm = merge(x=tm, y=aus, by.x="alter_examiner_id", by.y="examiner_id", all.x=TRUE)
tm = tm %>% rename(alter_art_unit=examiner_art_unit, alter_wg=wg)
tm = drop_na(tm)

egoNodes = subset(tm, select=c(ego_examiner_id,ego_art_unit, ego_wg)) %>% rename(examiner_id=ego_examiner_id,art_unit=ego_art_unit,wg=ego_wg)
alterNodes = subset(tm, select=c(alter_examiner_id,alter_art_unit, alter_wg))%>% rename(examiner_id=alter_examiner_id,art_unit=alter_art_unit,wg=alter_wg)
nodes = rbind(egoNodes, alterNodes)
nodes = distinct(nodes)
nodes = nodes %>% group_by(examiner_id) %>% summarise(examiner_id=first(examiner_id), art_unit=first(art_unit), wg=first(wg))

net = graph_from_data_frame(d=tm, vertices=nodes, directed=TRUE)
net
```

```{r}
# centrality
Degree <- degree(net, v=V(net))
Betweenness <- betweenness(net)
Eigenvector <- evcent(net)$vector

V(net)$size = Degree
V(net)$eig = round(Eigenvector,2)
V(net)$bet = round(Betweenness,2)

V(net)$color = nodes$art_unit

centralities <- cbind(Degree, Eigenvector, Betweenness)
centralities = round(centralities,2)
centralities = data.frame(centralities)

V(net)
```

# Graph

```{r}
ggraph(net, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=size, color=color), show.legend=T)
```

It looks like there are examiners leaving their group and seeking advice elsewhere (those are the smaller groups formed around the larger two groups).

It also looks like in the blue network, there are some nodes in black corresponding to examiners starting with identifier 16. Let's look at id 64507 and get a summary.

```{r}
exam <- app %>% filter(examiner_id==64507)
summary(exam)
exam
```